Is structure based drug design ready for selectivity optimization?

### 2018-10-28 manuscript.tex
updated figure 1

\paragraph{Assessing the ability of alchemical free energy methods to predict selectivity}
Here, we investigate the utility of alchemical free energy calculations for the prediction of selectivity, hereafter taken to mean the $\Delta$$\Delta$G in binding free energies of the same compound for two targets. We employed state of the art relative free energy calculations~\citep{Wang:J.Am.Chem.Soc.:2015} to predict the selectivities of two different congeneric ligand series~\citep{Shao2013-oe, Blake2016-su}, as well as present a statistical reanalysis of absolute free energy calculations on three different bromodomain inhibitors~\citep{Abel2017-gw}. \todo{Insert summary of results}.

@@

\section{Results}
\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure1.pdf}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-1}
{\bf Absolute free energy calculations show a wide range of errors in predicting selectivity} \\
({\bf A})  A representative structure of BRD2 (4ALG) bound to ligand RVX-OH, highlighting the binding mode of the ligands considered in this paper. 
({\bf B}) The chemical structure of all three brodomain inhibitors used in this reference. 
({\bf C}) he bromodomain family tree, adapted from Aldeghi et al., 2016. Bromodomains shown in green (from original) had absolute free energy calculations performed on them. The orange box highlights the bromodomains for which RVX-208, RXV-OH and Bromosporine calculations were performed. The blue box contains bromodomains for which only Bromosporine calculations were performed. 
({\bf D})The calculated $\Delta\Delta$G for each possible pairing of bromodomains. Orange ellipses represent a pair in which one or both of the bromodomains are from Family II (orange box, 1C). The blue ellipses contain only bromodomains from outside family II (blue box, 1C). The radii of the ellipses show the statistical uncertainty (in the Y-axis) or the experimental uncertainty (in the X-axis). The solid black line corresponds to absolute agreement between predicted $\Delta\Delta$G (Y-axis) and experiment $\Delta\Delta$G (X-axis). The Mean Unsigned Error (MUE) and Root Mean Square Error (RMSE) were calculated and shown with bootstrapped 95$\%$ confidence intervals, as described in the methods section.
}
\end{fullwidth}
\end{figure}

@@

\todo{figure out why figure4 is rendering this way}
\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure4.png}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-4}
{\bf Relative free energy calculations can accurately predict potency, but show larger errors for selectivity predictions.} \\
({\bf A}) Single target potencies for CDK2, ERK2 and CDK9 from both the Shao and Blake datasets, with experimental values on the X-axis and calculated values on the Y-axis. Each black point corresponds to a ligand for a given target. All values are shown in units of kcal/mol. The horizontal error bars show the assumed experimental uncertainty of 0.3 kcal/mol\citep{BROWN2009420}. To better highlight outliers that are unlikely due simply to forcefield errors, we presume the forcefield error ($\sigma_\mathrm{FF} \approx$ 0.9 kcal mol$^{-1}$~\cite{Harder:J.Chem.TheoryComput.:2016}) also behaves as a random error. We show the total estimated statistical and forcefield error ($\sqrt{\sigma_\mathrm{FF}^2 + \sigma_\mathrm{calc}^2}$) as vertical error bars. The black line indicates agreement between calculation and experiment, while the gray shaded region represent 1.36 kcal/mol (or 1 log unit) error. The MUE and RMSE are shown on each plot with bootstrapped 95$\%$ confidence intervals.
({\bf B}) The indicated $\Delta\Delta$G is shown for each ligand. 
}
\end{fullwidth}
\end{figure}

>>>>

\paragraph{Assessing the ability of alchemical free energy methods to predict selectivity}
Here, we investigate the utility of alchemical free energy calculations for the prediction of selectivity, hereafter taken to mean the $\Delta \Delta$G in binding free energies of the same compound for two targets. We employed state of the art relative free energy calculations~\citep{Wang:J.Am.Chem.Soc.:2015} to predict the selectivities of two different congeneric ligand series~\citep{Shao2013-oe, Blake2016-su}, as well as present a statistical reanalysis of absolute free energy calculations on three different bromodomain inhibitors~\citep{Abel2017-gw}. \todo{Insert summary of results}.

@@

\section{Results}
\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure1.png}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-1}
{\bf Free energy calculations speed up selectivity optimization} \\
({\bf A})  The effect of correlation on expected errors for predicting selectivity ($\sigma_{selectivity}$) in kcal/mol. Each curve represents a different combination of target errors ($\sigma_1$ and $\sigma_2$). 
({\bf B}) The change in selectivity for molecules proposed by medicinal chemists optimizing a lead candidate can be modeled by a normal distribution centered on 0 with a standard deviation of 1 kcal/mol (black curve). Each green curve corresponds to the distribution of compounds made after screening for a 1 log unit (1.4 kcal/mol) improvement in selectivity with a free energy methodology with a 1 kcal/mol per target error and a particular correlation. The shade region of each curve corresponds to the compounds with a real 1 log unit  improvement in selectivity. The speed up is calculated as the ratio of the percentage of compounds made with a real 1 log unit improvement to the percentage of compounds that would be expected in the original distribution.  
({\bf C}) The speedup (y-axis, log scale) expected for 100x (2 log units, 2.8 kcal/mol) selectivity optimization as a function of correlation coefficient $\rho$. Each curve corresponds to a different $\sigma_{target}$ value.  
}
\end{fullwidth}
\end{figure}

@@

\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure4.png}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-4}
{\bf Relative free energy calculations can accurately predict potency, but show larger errors for selectivity predictions.} \\
({\bf A}) Single target potencies for CDK2, ERK2 and CDK9 from both the Shao and Blake datasets, with experimental values on the X-axis and calculated values on the Y-axis. Each black point corresponds to a ligand for a given target. All values are shown in units of kcal/mol. The horizontal error bars show the assumed experimental uncertainty of 0.3 kcal/mol\citep{BROWN2009420}. To better highlight outliers that are unlikely due simply to forcefield errors, we presume the forcefield error ($\sigma_\mathrm{FF} \approx$ 0.9 kcal mol$^{-1}$~\cite{Harder:J.Chem.TheoryComput.:2016}) also behaves as a random error. We show the total estimated statistical and forcefield error ($\sqrt{\sigma_\mathrm{FF}^2 + \sigma_\mathrm{calc}^2}$) as vertical error bars. The black line indicates agreement between calculation and experiment, while the gray shaded region represent 1.36 kcal/mol (or 1 log unit) error. The MUE and RMSE are shown on each plot with bootstrapped 95$\%$ confidence intervals.
({\bf B}) The indicated $\Delta\Delta$G is shown for each ligand. 
}
\end{fullwidth}
\end{figure}
